---
title: "Reproducible Research - Project2"
author: "Stanislas de Bazelaire"
date: "2/21/2021"
output:
  html_document: default
  pdf_document: default
---

# Introduction

This project involves exploring the U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database. This database tracks characteristics of **major storms and weather events in the United States**, including when and where they occur, as well as estimates of any fatalities, injuries, and property damage.

We focus on two types of storms damage: 

1. **Human**, by looking at injuries and deaths;

2. **Economic**, by looking at damage to properties and crops.

Our goal is to identify the most damaging types of storms in the US. We narrow the list of most damaging storms to the **top five**.

# I. Data processing

## A. Loading the data

Let's start by loading the data and caching this step to avoid having to reload the data each time we run the script:
```{r, echo=TRUE,results='hide',cache=TRUE}
setwd("C:/Users/stanislas_bazelaire/Box/Stanislas_Bazelaire/Data science/Reproducible research")
dir.create("Stormdata")
library(downloader)
download("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2","./Stormdata/stormdata.csv.bz2",mode="wb") 
data<-read.csv("./Stormdata/stormdata.csv.bz2",header=TRUE,sep=",")
```

## B. Injuries
Let's compute the sum of injuries by storm event type with the aggregate function and then:

-keep events only for which the number of injured is greater than zero

-order the events by number of injuries

-keep only the top 5 events

```{r, echo=TRUE,results='hide'}
injuries<-aggregate(data$INJURIES~data$EVTYPE,data=data,sum)
names(injuries)<-c("event","injured")
injuries<-injuries[injuries$injured>0,]
injuries<-injuries[order(-injuries$injured),]
top5injuries<-injuries[1:5,]
```

## C. Deaths
Let's proceed exactly in the same way for the number of deaths:
```{r, echo=TRUE,results='hide'}
fat<-aggregate(data$FATALITIES~data$EVTYPE,data=data,sum)
names(fat)<-c("event","deaths")
fat<-fat[fat$deaths>0,]
fat<-fat[order(-fat$deaths),]
top5fat<-fat[1:5,]
```

## D. The issue of units for economic damage

Units are indicated in two colmuns: **PROPDMGEXP** and **CROPDMGEXP**.

Let's map the symbols of these two columns as per the table below. To do this we create a vector ('unit') which contains all the different units and a vector ('v') which contains all the values we want to assign to the different units:

```{r, echo=TRUE}
allunits<-data.frame(c(data$PROPDMGEXP,data$CROPDMGEXP))
units<-unique(allunits)
v<-c(10^3,10^6,1,10^9,10^6,1,1,10^5,10^6,1,10^4,10^2,10^3,10^2,10^7,10^2,1,1,10^8,10^3)
values<-data.frame(units,v)
names(values)<-c("UNIT","VALUE")
print(values)
```

## E. Property damage

We proceed as follows:

1. We create a dataframe with three columns: the storm type, the damage, the unit

2. We merge the dataframe with the units dataframe above

3. We compute the amount of the damage by adding a new column equal to the damage times the unit

4. We compute total damages per type of storm event (aggregate function) and (as before) narrow the results to the top 5

```{r, echo=TRUE}
prop<-data.frame(data$EVTYPE,data$PROPDMG,data$PROPDMGEXP)
names(prop)<-c("EVTYPE","PROPDMG","UNIT")
prop2<-merge(prop,values,by.x = "UNIT",by.y = "UNIT",all.x = TRUE, all.y = FALSE )
prop2$DMG<-prop2$PROPDMG*prop2$VALUE
propdmg<-aggregate(prop2$DMG~prop2$EVTYPE,data=prop2,sum)
names(propdmg)<-c("EVTYPE","DMG")
propdmg<-propdmg[order(-propdmg$DMG),]
propdmg<-propdmg[1:5,]
propdmg$DMG<-propdmg$DMG/10^9
propdmg$DMG<-round(propdmg$DMG,digits = 1)
```

## F. Crops damage

We do exactly the same as for property damages

```{r, echo=TRUE}
crop<-data.frame(data$EVTYPE,data$CROPDMG,data$CROPDMGEXP)
names(crop)<-c("EVTYPE","CROPDMG","UNIT")
crop2<-merge(crop,values,by.x = "UNIT",by.y = "UNIT",all.x = TRUE, all.y = FALSE )
crop2$DMG<-crop2$CROPDMG*prop2$VALUE
cropdmg<-aggregate(crop2$DMG~crop2$EVTYPE,data=crop2,sum)
names(cropdmg)<-c("EVTYPE","DMG")
cropdmg<-cropdmg[order(-cropdmg$DMG),]
cropdmg<-cropdmg[1:5,]
cropdmg$DMG<-cropdmg$DMG/10^9
cropdmg$DMG<-round(cropdmg$DMG,digits = 1)
```

# II. Results

## A. Human cost

We will plot the cost of storm events in terms of injuries and deaths using the ggplot2 package and aggregate the two charts (p1 and p2 below) in one chart using the ggpubr package:

```{r,echo=TRUE,fig.height=8,fig.width=12}
library(ggplot2)
library(ggpubr)

p1<-ggplot(data=top5injuries,aes(x=reorder(event,-injured),y=injured))+geom_bar(stat="identity",colour="blue",fill="blue")+xlab("")+ylab("number injured")+geom_text(aes(label=injured),position=position_dodge(width=0.9),vjust=-0.25)

p2<-ggplot(data=top5fat,aes(x=reorder(event,-deaths),y=deaths))+geom_bar(stat="identity",colour="red",fill="red")+ylab("number of deaths")+xlab("")+geom_text(aes(label=deaths),position=position_dodge(width=0.9),vjust=-0.25)

ggarrange(p1, p2, ncol = 2, nrow = 1)
```

## B. Economic cost

We proceed exactly as above:

```{r, echo=TRUE,fig.height=8,fig.width=12}
library(ggplot2)
library(ggpubr)

q1<-ggplot(data=propdmg,aes(x=reorder(EVTYPE,-DMG),y=DMG))+geom_bar(stat="identity",colour="red",fill="red")+ylab("property damage $ bn")+xlab("")+geom_text(aes(label=DMG),position=position_dodge(width=0.9),vjust=-0.25)

q2<-ggplot(data=cropdmg,aes(x=reorder(EVTYPE,-DMG),y=DMG))+geom_bar(stat="identity",colour="blue",fill="blue")+ylab("crops damage $ bn")+xlab("")+geom_text(aes(label=DMG),position=position_dodge(width=0.9),vjust=-0.25)

ggarrange(q1, q2, ncol = 2, nrow = 1)
```
